\papertitle

### 2021-05-24 source_separation/paper/paper.tex
Small changes from SMC reviews

	Source separation is an important addition to improve orchestration as it allows for the orchestration of more complex sounds in a way that makes sense in an orchestral context. Orchestral music often uses multiple layers, so by separating the layers of a target and individually orchestrating them, the resulting solution is improved.

@@

	\begin{figure}[t]
		\centering
			\includegraphics[width=\columnwidth]{figures/diagram.png}
			\caption{Diagram comparing the process of orchestration with and without separation. At the top, the full target orchestration is created simply by orchestrating the target. At the bottom, the separation process is performed, and the resulting sub-targets are individually orchestrated with subsets of the orchestra before being recombined to obtain the full orchestration.}\label{fig:full_diagram}
	\end{figure}

@@

		\begin{enumerate}
			\item Static sounds in which the harmonic average across time is a fitting representation of the sound
			\item Sounds in which there is at least some pitched content and not only noise
		\end{enumerate}			
		The sounds chosen include alarms, bells, engine noises, soundscapes, synthesizer chords, and sound effects. Each target that we used for testing was a combination of four randomly chosen source sounds from the group of 90 sounds. 
		When a target is created, the longest of the four sources is chosen to start playing at the beginning of the target. The other three sources are then randomly assigned different times to begin playing such that they will start between the beginning of the target and half-way through the longest source sound.

		\subsection{Separation Methods}
	We consider source separation models trained to tackle two distinct separation problems. The first is music source separation (Open-Unmix, Demucs, ConvTasnet). For these architectures, we used models that were pre-trained on the MUSDB18 dataset~\cite{MUSDB18}. Thus, these models output 4 stereo tracks that correspond to broad instrumental categories defined in the SiSEC 2018 Mus evaluation campaign: \textit{vocals, drums, bass} and \textit{other}~\cite{Stoter_SiSEC}. Although the audio sources we consider in this work are often non-musical, we hope that these source separation methods can still improve the orchestration by separating other types of sound in a musically meaningful way.
	The second type of model we tested are ones trained for universal sound separation (TDCN++, NMF), which \cite{tdcnpp} defines as \textit{separating mixtures of arbitrary sounds of different types}.

@@

			The \textbf{Demucs}~\cite{demucs} architecture is a deep learning model for musical source separation. It takes as input the stereo mixture, and outputs 4 stereo waveforms corresponding to the four categories: vocals, bass, drums and other. Demucs is based on a convolutional encoder, a bidirectional LSTM, and a convolutional decoder. The encoder and the decoder are linked with skip U-Net connections. Both the encoder and the decoder are made of 6 blocks, each made of a combination of a convolutional layer with kernel size 8, followed by ReLU activation. Finally there is another convolution with kernel size 1, followed by gated linear units (GLU). The U-network structure allows the decoder to directly access the input signal and to transfer its phase, improving results.

@@

		\subsection{Testing}
		In order to compare the effectiveness of the separation methods, a full target orchestration and a ground truth orchestration are done for each target. The full target orchestration is the orchestration of the target without any separation performed. The ground truth orchestration takes the four sources that make up the target and orchestrates each one separately, then combines them. This ground truth represents the orchestration that could take place if the separation method was perfect and separated a target exactly into its constituent sources.

		The procedure we use for testing is as follows:
		\begin{enumerate}
			\item The target is created as a combination of four randomly chosen sources, which are offset to begin playing at different times
			\item The full target, without any separation performed, is orchestrated using the entire orchestra
			\item Each separation method is performed on the target, each splitting the target into four sub-targets 
			\item The full orchestra is randomly split into four equal sized subsets
			\item For each method, the four sub-targets are separately orchestrated, each using a different subset of the orchestra, and then each orchestration is combined to play simultaneously
			\item The ground truth orchestration is created by orchestrating the four sources, using the same orchestra splits from step 4, and then combined to play simultaneously
			\item The distances between target and orchestration are calculated for the full target orchestration, the separated orchestration, and the ground truth orchestration 
		\end{enumerate}

		\begin{figure}[t]
			\centering
				\includegraphics[width=\columnwidth]{figures/evaluation_diagram.png}
				\caption{Diagram showing the testing process. Note that in both orchestrations that use a split orchestra, the split is the same.}\label{fig:eval_diagram}
		\end{figure}

@@

		The distance metric cuts the target and solution into successive frames that are 4,000 samples in length, which is approximately 90 milliseconds at a sampling rate of 44.1 kHz. The spectral distance, as defined in Eqn. \ref{eqn:distance}, is calculated between corresponding frames of the target and solution. The spectral distance metric is proposed in \cite{Cella2020} as part of the cost function used in Orchidea during the optimization, and used in \cite{Cella2020b} to compare accuracies of orchestrations. The equation takes in the full FFT of the target $x$ and full FFT of the solution $\tilde{x}$. Then for each bin $k$ of the FFT, it calculates the absolute difference between the values. The differing values of $\lambda_1$ and $\lambda_2$ allow the metric to penalize the solution in different ways. For our purposes, we used $\lambda_1 = 0.5$ and $\lambda_2 = 10$, which penalizes a solution that overshoots the harmonic energy of the target.

@@

	\begin{table}[t]
		\begin{center}
			\begin{tabular}{|c|c|}
				\hline
				& Average distance \\
				\hline
				Demucs & 25.11 \\
				\hline
				Open-Unmix & 24.48\\
				\hline
				TDCN & 26.44\\
				\hline
				NMF & 22.44 \\
				\hline
				TDCN++ & 24.38 \\
				\hline
				Full target & 24.96 \\
				\hline
				Ground truth & 24.19 \\
				\hline
			\end{tabular}
		\end{center}
		\caption{Average across 300 targets showing the distance between target and orchestration for various methods. ``Full target'' means no separation was performed. Lower values are better.}
		\label{tab:distances}
	\end{table}

>>>>

	Source separation is an important addition to improve orchestration as it allows for the orchestration of more complex sounds in a way that makes sense in an orchestral context. Orchestral music often uses multiple layers, so by separating the layers of a target and individually orchestrating them, the resulting orchestration may improve.

@@

	\begin{figure*}[t]
		\centering
			\includegraphics[width=\textwidth]{figures/diagram.png}
			\caption{Diagram comparing the process of orchestration with and without separation. At the top, the full target orchestration is created simply by orchestrating the target. At the bottom, the separation process is performed, and the resulting sub-targets are individually orchestrated with subsets of the orchestra before being recombined to obtain the full orchestration.}\label{fig:full_diagram}
	\end{figure*}

@@

		\begin{enumerate}
			\item Static sounds that do not change harmonically over time
			\item Sounds in which there is at least some pitched content and not only noise
		\end{enumerate}			
		The sounds chosen include alarms, bells, engine noises, soundscapes, synthesizer chords, and sound effects. Each target that we used for testing was a combination of four randomly chosen source sounds from the group of 90 sounds. 
		When a target is created, the longest of the four sources is chosen to start playing at the beginning of the target. The other three sources are then randomly assigned different times to begin playing such that they will start between the beginning of the target and half-way through the longest source sound.

		\subsection{Separation Methods}
	We consider source separation models trained to tackle two distinct separation problems. The first is music source separation (Open-Unmix, Demucs, ConvTasnet). For these architectures, we used models that were pre-trained on the MUSDB18 dataset~\cite{MUSDB18}. Thus, these models output 4 stereo tracks that correspond to broad instrumental categories defined in the SiSEC 2018 Mus evaluation campaign: \textit{vocals, drums, bass} and \textit{other}~\cite{Stoter_SiSEC}. It is important here to make the distinction that we did not train these models ourselves. Therefore, we cannot guarantee that the data the models were trained on is optimal for solving our problem. However, one of the goals of this paper is to explore whether pre-trained music source separation models can separate non-musical sounds in a way that improves orchestration.
	The second type of model we tested are ones that perform universal sound separation (TDCN++, NMF), which \cite{tdcnpp} defines as \textit{separating mixtures of arbitrary sounds of different types}. These methods more closely match our problem of separating non-musical target sounds.

@@

			The \textbf{Demucs}~\cite{demucs} architecture is a deep learning model for musical source separation. It takes the stereo mixture as input, and outputs 4 stereo waveforms corresponding to the four categories: vocals, bass, drums and other. Demucs is based on a convolutional encoder, a bidirectional LSTM, and a convolutional decoder. The encoder and the decoder are linked with skip U-Net connections. Both the encoder and the decoder are made of 6 blocks, each made of a combination of a convolutional layer with kernel size 8, followed by ReLU activation. Finally there is another convolution with kernel size 1, followed by gated linear units (GLU). The U-network structure allows the decoder to directly access the input signal and to transfer its phase, improving results.

@@

		\subsection{Testing}\label{subsec:testing}
		In order to compare the effectiveness of the separation methods, a full target orchestration and a ground truth orchestration are done for each target. The full target orchestration is the orchestration of the target without any separation performed. The ground truth orchestration takes the four sources that make up the target and orchestrates each one separately, then combines them. This ground truth represents the orchestration that could take place if the separation method was perfect and separated a target exactly into its constituent sources.

		The procedure we use for testing is as follows:
		\begin{enumerate}
			\item The target is created as a combination of four randomly chosen sources, which are offset to begin playing at different times.
			\item The full target, without any separation performed, is orchestrated using the entire orchestra. This is called the \textbf{full target orchestration}.
			\item Each separation method is performed on the target, each splitting the target into four sub-targets. 
			\item The full orchestra is randomly split into four equal sized subsets, each containing 7 to 8 instruments.
			\item For each method, the four sub-targets are separately orchestrated, each using a different subset of the orchestra, and then each orchestration is combined to play simultaneously. This creates the \textbf{separated orchestrations}.
			\item The \textbf{ground truth orchestration} is created by orchestrating the four sources, using the same orchestra subsets from step 4, and then combined to play simultaneously.
			\item The distances between target and orchestration are calculated for the full target orchestration, the separated orchestration, and the ground truth orchestration. 
		\end{enumerate}

		\begin{figure*}[t]
			\centering
				\includegraphics[width=\textwidth]{figures/evaluation_diagram.png}
				\caption{Diagram showing the testing process. Note that in both orchestrations that use a split orchestra, they use the same division of the orchestra to orchestrate their four sub-targets.}\label{fig:eval_diagram}
		\end{figure*}

@@

		The distance metric cuts the target and solution into successive frames that are 4,000 samples in length, which is approximately 90 milliseconds at a sampling rate of 44.1 kHz. The spectral distance, as defined in Eqn. \ref{eqn:distance}, is calculated between corresponding frames of the target and solution. The spectral distance metric is proposed in \cite{Cella2020} as part of the cost function used in Orchidea during the optimization, and used in \cite{Cella2020b} to compare accuracies of orchestrations. The equation takes in the FFT of the target $x$ and FFT of the solution $\tilde{x}$. Then for each bin $k$ of the FFT, it calculates the absolute difference between the values. The differing values of $\lambda_1$ and $\lambda_2$ allow the metric to penalize the solution in different ways. For our purposes, we used $\lambda_1 = 0.5$ and $\lambda_2 = 10$, which penalizes a solution that overshoots the harmonic energy of the target.

@@

	\begin{table}[t]
		\begin{center}
			\begin{tabular}{|c|c|}
				\hline
				& Average distance \\
				\hline
				Demucs & 25.11 \\
				\hline
				Open-Unmix & 24.48\\
				\hline
				TDCN & 26.44\\
				\hline
				NMF & 22.44 \\
				\hline
				TDCN++ & 24.38 \\
				\hline
				Full target orchestration & 24.96 \\
				\hline
				Ground truth orchestration & 24.19 \\
				\hline
			\end{tabular}
		\end{center}
		\caption{Average across 300 targets showing the distance between target and orchestration for various methods. For Demucs, Open-Unmix, TDCN, NMF, and TDCN++, the value shown is the average distance between the target sound and the ``separated orchestration'' created from the output of the given separation method (see Sec. \ref{subsec:testing}). Similarly, ``Full target orchestration'' and ``Ground truth orchestration'', as defined in Sec. \ref{subsec:testing}, are the average values between target and orchestrated solution. Lower values are better.}
		\label{tab:distances}
	\end{table}
